---
title: "Preliminary ELSA Machine Learnings Tests"
author: "Jonathan Seiden"
date: "8/18/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readstata13)
library(tidyverse)
library(glmnet)
library(caret)
library(fastDummies)
```

# Data consolidation process  

In this section, we process the data to get it into a format where each row is a child. 

## Child and Teacher Observation  
First we input the child and teacher observations and process them.

For each child in the COP data, we calculate:

1) The average for the child for each of the indicators across sweeps
2) The class average for each indicator *omitting* the child her/himself
3) The class standard deviation for each indicator *omitting* the child her/himself (only including children with 10 or more sweeps)

We then calculate the class average of the TOP indicators for the adults in the class by averaging across sweeps, and merge this data (one to many) with the child-level data. This merged data set contains XXX children in XXX classes

```{r y1_obs, warning=FALSE}
#Input the Year One long child and teacher observation data 
y1_child_obs_raw <- read.dta13("y1o_c_long.dta")
y1_teacher_obs_raw <- read.dta13("y1o_t_long.dta")
y1_coverpage_obs <- read.dta13("y1o_coverpage.dta")

#Re-format the child data so that it is one row per child
y1_child_obs <- y1_child_obs_raw %>%
  mutate(cid = ifelse(childid == "N/A", o_c_uniqueid, childid)) %>%
  group_by(cid, classid) %>%
  mutate(nsweeps = n()) %>%
  mutate_at(vars(o_c_verbal:o_c_focus), as.character) %>% 
  select(c(classid, nsweeps, o_c_verbal:o_c_focus)) %>%
  dummy_cols(select_columns = c("o_c_verbal", "o_c_towhom", "o_c_schedule", "o_c_interaction", "o_c_typetask", "o_c_involvement", "o_c_focus"),
             remove_selected_columns = TRUE) %>% 
  group_by(cid, classid) %>%
  summarize(across(everything(), mean, na.rm = TRUE)) %>%
  filter(nsweeps >= 10 ) %>% #THIS IS AN ARBITRARY PARAMETER
  group_by(classid) %>%
  mutate(nclass = n(),
         #This horrid mess calculated the leave-out mean in the class for each student for each COP variable
         `o_c_verbal_Fuss/Cry (FC)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_verbal_Fuss/Cry (FC)`[-.x]), na.rm = TRUE),
         `o_c_verbal_Listening (L)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_verbal_Listening (L)`[-.x]), na.rm = TRUE),
         `o_c_verbal_No (N)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_verbal_No (N)`[-.x]), na.rm = TRUE),
         `o_c_verbal_Talk (T)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_verbal_Talk (T)`[-.x]), na.rm = TRUE),
         `o_c_verbal_Talk, Sounds (TS)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_verbal_Talk, Sounds (TS)`[-.x]), na.rm = TRUE),
         `o_c_verbal_NA_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_verbal_NA`[-.x]), na.rm = TRUE),
         `o_c_towhom_Child (C): One other child_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_towhom_Child (C): One other child`[-.x]), na.rm = TRUE),
         `o_c_towhom_No talk/listen (NT)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_towhom_No talk/listen (NT)`[-.x]), na.rm = TRUE),
         `o_c_towhom_Self (S)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_towhom_Self (S)`[-.x]), na.rm = TRUE),
         `o_c_towhom_Small Group, No Teacher (SG)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_towhom_Small Group, No Teacher (SG)`[-.x]), na.rm = TRUE),
         `o_c_towhom_Small Group, With Teacher (SGT)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_towhom_Small Group, With Teacher (SGT)`[-.x]), na.rm = TRUE),
         `o_c_towhom_Teacher (T): Teacher or assistant (adult)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_towhom_Teacher (T): Teacher or assistant (adult)`[-.x]), na.rm = TRUE),
         `o_c_towhom_Whole Group, No Teacher (WG)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_towhom_Whole Group, No Teacher (WG)`[-.x]), na.rm = TRUE),
         `o_c_towhom_Whole Group, With Teacher (WGT)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_towhom_Whole Group, With Teacher (WGT)`[-.x]), na.rm = TRUE),
         `o_c_towhom_NA_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_towhom_NA`[-.x]), na.rm = TRUE),
         `o_c_schedule_Centers (C)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Centers (C)`[-.x]), na.rm = TRUE),
         `o_c_schedule_Combination of Small Group and Centers (SGC)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Combination of Small Group and Centers (SGC)`[-.x]), na.rm = TRUE),
         `o_c_schedule_Gym (G)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Gym (G)`[-.x]), na.rm = TRUE),
         `o_c_schedule_IAT_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_IAT`[-.x]), na.rm = TRUE),
         `o_c_schedule_Meal Time (MT)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Meal Time (MT)`[-.x]), na.rm = TRUE),
         `o_c_schedule_Nap (N)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Nap (N)`[-.x]), na.rm = TRUE),
         `o_c_schedule_Playground (P)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Playground (P)`[-.x]), na.rm = TRUE),
         `o_c_schedule_Small Groups (SG)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Small Groups (SG)`[-.x]), na.rm = TRUE),
         `o_c_schedule_Special (SP)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Special (SP)`[-.x]), na.rm = TRUE),
         `o_c_schedule_Transition (T)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Transition (T)`[-.x]), na.rm = TRUE),
         `o_c_schedule_Whole Group (WG)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Whole Group (WG)`[-.x]), na.rm = TRUE),
         `o_c_interaction_Alone (AL)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_Alone (AL)`[-.x]), na.rm = TRUE),
         `o_c_interaction_Associative (AS)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_Associative (AS)`[-.x]), na.rm = TRUE),
         `o_c_interaction_Cooperative (C)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_Cooperative (C)`[-.x]), na.rm = TRUE),
         `o_c_interaction_Non-Academic (NA)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_Non-Academic (NA)`[-.x]), na.rm = TRUE),
         `o_c_interaction_Onlooker (ON)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_Onlooker (ON)`[-.x]), na.rm = TRUE),
         `o_c_interaction_Parallel (P)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_Parallel (P)`[-.x]), na.rm = TRUE),
         `o_c_interaction_Social (SOC)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_Social (SOC)`[-.x]), na.rm = TRUE),
         `o_c_interaction_Time Out (TO)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_Time Out (TO)`[-.x]), na.rm = TRUE),
         `o_c_interaction_Unoccupied (UN)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_Unoccupied (UN)`[-.x]), na.rm = TRUE),
         `o_c_interaction_NA_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_NA`[-.x]), na.rm = TRUE),
         `o_c_typetask_Disruptive (D)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_Disruptive (D)`[-.x]), na.rm = TRUE),
         `o_c_typetask_Non-Sequential (NS)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_Non-Sequential (NS)`[-.x]), na.rm = TRUE),
         `o_c_typetask_None (N)"_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_None (N)`[-.x]), na.rm = TRUE),
         `o_c_typetask_Other (O)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_Other (O)`[-.x]), na.rm = TRUE),
         `o_c_typetask_Passive Instruction (PI)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_Passive Instruction (PI)`[-.x]), na.rm = TRUE),
         `o_c_typetask_Sequential (SQ)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_Sequential (SQ)`[-.x]), na.rm = TRUE),
         `o_c_typetask_Sleep (S)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_Sleep (S)`[-.x]), na.rm = TRUE),
         `o_c_typetask_Social (SOC)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_Social (SOC)`[-.x]), na.rm = TRUE),
         `o_c_typetask_Time Out (TO)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_Time Out (TO)`[-.x]), na.rm = TRUE),
         `o_c_typetask_NA_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_NA`[-.x]), na.rm = TRUE),
         `o_c_involvement_High (H)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_involvement_High (H)`[-.x]), na.rm = TRUE),
         `o_c_involvement_Low (L)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_involvement_Low (L)`[-.x]), na.rm = TRUE),
         `o_c_involvement_Medium (M)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_involvement_Medium (M)`[-.x]), na.rm = TRUE),
         `o_c_involvement_Medium High (MH)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_involvement_Medium High (MH)`[-.x]), na.rm = TRUE),
         `o_c_involvement_Medium Low (ML)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_involvement_Medium Low (ML)`[-.x]), na.rm = TRUE),
         `o_c_involvement_NA_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_involvement_NA`[-.x]), na.rm = TRUE),
         `o_c_focus_Art (A)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Art (A)`[-.x]), na.rm = TRUE),
         `o_c_focus_Drama (D)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Drama (D)`[-.x]), na.rm = TRUE),
         `o_c_focus_Gross Motor (GM)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Gross Motor (GM)`[-.x]), na.rm = TRUE),
         `o_c_focus_Language Arts - Writing (LAW)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Language Arts - Writing (LAW)`[-.x]), na.rm = TRUE),
         `o_c_focus_Literacy - Writing (LW)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Literacy - Writing (LW)`[-.x]), na.rm = TRUE),
         `o_c_focus_Literacy (L)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Literacy (L)`[-.x]), na.rm = TRUE),
         `o_c_focus_Math (M)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Math (M)`[-.x]), na.rm = TRUE),
         `o_c_focus_Music and Movement (MM)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Music and Movement (MM)`[-.x]), na.rm = TRUE),
         `o_c_focus_None (N)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_None (N)`[-.x]), na.rm = TRUE),
         `o_c_focus_Other (O)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Other (O)`[-.x]), na.rm = TRUE),
         `o_c_focus_Other Language (OL)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Other Language (OL)`[-.x]), na.rm = TRUE),
         `o_c_focus_Reading (R)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Reading (R)`[-.x]), na.rm = TRUE),
         `o_c_focus_Science (SC)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Science (SC)`[-.x]), na.rm = TRUE),
         `o_c_focus_Social Studies (SS)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Social Studies (SS)`[-.x]), na.rm = TRUE),
         `o_c_focus_NA_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_NA`[-.x]), na.rm = TRUE)) %>%
  ungroup 

#Re-format the teacher data so that it is one row per class
y1_teacher_obs <- y1_teacher_obs_raw %>% 
  dummy_cols(select_columns = c("o_t_verbal_o", "o_t_whom_o", "o_t_schedule_o", 
                                "o_t_task_o", "o_t_instruct", "o_t_focus_o", 
                                "o_t_tone_o", "o_t_attention_o", "o_t_es_o"),
             remove_selected_columns = TRUE) %>% 
    group_by(classid) %>%
  summarize(
    nsweeps = n(),
    nadult = length(unique(o_t_uniqueid)),
    across(starts_with(c("o_t_verbal_o", "o_t_whom_o", "o_t_schedule_o", 
                                "o_t_task_o", "o_t_instruct", "o_t_focus_o", 
                                "o_t_tone_o", "o_t_attention_o", "o_t_es_o")),
           mean)) %>%
  select(-ends_with("_")) %>% 
  ungroup

#Merge teacher and child observations 
y1_obs <- left_join(y1_child_obs, y1_teacher_obs, by = "classid")

```

```{r y2_obs, warning=FALSE}
#Input the Year One long child and teacher observation data 
y2_child_obs_raw <- read.dta13("y2o_c_long.dta")
y2_teacher_obs_raw <- read.dta13("y2o_t_long.dta")
y2_coverpage_obs <- read.dta13("y2o_coverpage.dta")

mean(y2_coverpage_obs$classid %in% y1_coverpage_obs$classid)

#Re-format the child data so that it is one row per child
y2_child_obs <- y2_child_obs_raw %>%
  group_by(cid, classid) %>%
  mutate(nsweeps = n()) %>%
  mutate_at(vars(o_c_verbal:o_c_focus), as.character) %>% 
  select(c(classid, nsweeps, o_c_verbal:o_c_focus)) %>%
  dummy_cols(select_columns = c("o_c_verbal", "o_c_towhom", "o_c_schedule", "o_c_interaction", "o_c_typetask", "o_c_involvement", "o_c_focus"),
             remove_selected_columns = TRUE) %>% 
  group_by(cid, classid) %>%
  summarize(across(everything(), mean, na.rm = TRUE)) %>%
  filter(nsweeps >= 10 ) %>% #THIS IS AN ARBITRARY PARAMETER
  group_by(classid) %>%
  mutate(nclass = n(),
         #This horrid mess calculated the leave-out mean in the class for each student for each COP variable
         `o_c_verbal_Fuss/Cry (FC)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_verbal_Fuss/Cry (FC)`[-.x]), na.rm = TRUE),
         `o_c_verbal_Listening (L)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_verbal_Listening (L)`[-.x]), na.rm = TRUE),
         `o_c_verbal_No (N)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_verbal_No (N)`[-.x]), na.rm = TRUE),
         `o_c_verbal_Talk (T)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_verbal_Talk (T)`[-.x]), na.rm = TRUE),
         `o_c_verbal_Talk, Sounds (TS)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_verbal_Talk, Sounds (TS)`[-.x]), na.rm = TRUE),
         `o_c_verbal_NA_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_verbal_NA`[-.x]), na.rm = TRUE),
         `o_c_towhom_Child (C): One other child_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_towhom_Child (C): One other child`[-.x]), na.rm = TRUE),
         `o_c_towhom_No talk/listen (NT)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_towhom_No talk/listen (NT)`[-.x]), na.rm = TRUE),
         `o_c_towhom_Self (S)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_towhom_Self (S)`[-.x]), na.rm = TRUE),
         `o_c_towhom_Small Group, No Teacher (SG)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_towhom_Small Group, No Teacher (SG)`[-.x]), na.rm = TRUE),
         `o_c_towhom_Small Group, With Teacher (SGT)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_towhom_Small Group, With Teacher (SGT)`[-.x]), na.rm = TRUE),
         `o_c_towhom_Teacher (T): Teacher or assistant (adult)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_towhom_Teacher (T): Teacher or assistant (adult)`[-.x]), na.rm = TRUE),
         `o_c_towhom_Whole Group, No Teacher (WG)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_towhom_Whole Group, No Teacher (WG)`[-.x]), na.rm = TRUE),
         `o_c_towhom_Whole Group, With Teacher (WGT)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_towhom_Whole Group, With Teacher (WGT)`[-.x]), na.rm = TRUE),
         `o_c_towhom_NA_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_towhom_NA`[-.x]), na.rm = TRUE),
         `o_c_schedule_Centers (C)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Centers (C)`[-.x]), na.rm = TRUE),
         `o_c_schedule_Combination of Small Group and Centers (SGC)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Combination of Small Group and Centers (SGC)`[-.x]), na.rm = TRUE),
         `o_c_schedule_Gym (G)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Gym (G)`[-.x]), na.rm = TRUE),
         `o_c_schedule_Individual activity w teacher (IAT)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Individual activity w teacher (IAT)`[-.x]), na.rm = TRUE),
         `o_c_schedule_Meal Time (MT)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Meal Time (MT)`[-.x]), na.rm = TRUE),
         `o_c_schedule_Nap (N)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Nap (N)`[-.x]), na.rm = TRUE),
         `o_c_schedule_Playground (P)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Playground (P)`[-.x]), na.rm = TRUE),
         `o_c_schedule_Small Groups (SG)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Small Groups (SG)`[-.x]), na.rm = TRUE),
         `o_c_schedule_Special (SP)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Special (SP)`[-.x]), na.rm = TRUE),
         `o_c_schedule_Transition (T)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Transition (T)`[-.x]), na.rm = TRUE),
         `o_c_schedule_Whole Group (WG)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_schedule_Whole Group (WG)`[-.x]), na.rm = TRUE),
         `o_c_interaction_Alone (AL)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_Alone (AL)`[-.x]), na.rm = TRUE),
         `o_c_interaction_Associative (AS)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_Associative (AS)`[-.x]), na.rm = TRUE),
         `o_c_interaction_Cooperative (C)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_Cooperative (C)`[-.x]), na.rm = TRUE),
         `o_c_interaction_Non-Academic (NA)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_Non-Academic (NA)`[-.x]), na.rm = TRUE),
         `o_c_interaction_Onlooker (ON)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_Onlooker (ON)`[-.x]), na.rm = TRUE),
         `o_c_interaction_Parallel (P)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_Parallel (P)`[-.x]), na.rm = TRUE),
         `o_c_interaction_Social (SOC)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_Social (SOC)`[-.x]), na.rm = TRUE),
         `o_c_interaction_Time Out (TO)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_Time Out (TO)`[-.x]), na.rm = TRUE),
         `o_c_interaction_Unoccupied (UN)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_Unoccupied (UN)`[-.x]), na.rm = TRUE),
         `o_c_interaction_NA_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_interaction_NA`[-.x]), na.rm = TRUE),
         `o_c_typetask_Disruptive (D)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_Disruptive (D)`[-.x]), na.rm = TRUE),
         `o_c_typetask_Non-Sequential (NS)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_Non-Sequential (NS)`[-.x]), na.rm = TRUE),
         `o_c_typetask_None (N)"_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_None (N)`[-.x]), na.rm = TRUE),
         `o_c_typetask_Other (O)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_Other (O)`[-.x]), na.rm = TRUE),
         `o_c_typetask_Passive Instruction (PI)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_Passive Instruction (PI)`[-.x]), na.rm = TRUE),
         `o_c_typetask_Sequential (SQ)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_Sequential (SQ)`[-.x]), na.rm = TRUE),
         `o_c_typetask_Sleep (S)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_Sleep (S)`[-.x]), na.rm = TRUE),
         `o_c_typetask_Social (SOC)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_Social (SOC)`[-.x]), na.rm = TRUE),
         `o_c_typetask_Time Out (TO)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_Time Out (TO)`[-.x]), na.rm = TRUE),
         `o_c_typetask_NA_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_typetask_NA`[-.x]), na.rm = TRUE),
         `o_c_involvement_High (H)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_involvement_High (H)`[-.x]), na.rm = TRUE),
         `o_c_involvement_Low (L)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_involvement_Low (L)`[-.x]), na.rm = TRUE),
         `o_c_involvement_Medium (M)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_involvement_Medium (M)`[-.x]), na.rm = TRUE),
         `o_c_involvement_Medium High (MH)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_involvement_Medium High (MH)`[-.x]), na.rm = TRUE),
         `o_c_involvement_Medium Low (ML)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_involvement_Medium Low (ML)`[-.x]), na.rm = TRUE),
         `o_c_involvement_NA_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_involvement_NA`[-.x]), na.rm = TRUE),
         `o_c_focus_Art (A)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Art (A)`[-.x]), na.rm = TRUE),
         `o_c_focus_Drama (D)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Drama (D)`[-.x]), na.rm = TRUE),
         `o_c_focus_Gross Motor (GM)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Gross Motor (GM)`[-.x]), na.rm = TRUE),
         `o_c_focus_Language Arts - Writing (LAW)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Language Arts - Writing (LAW)`[-.x]), na.rm = TRUE),
         `o_c_focus_Literacy - Writing (LW)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Literacy - Writing (LW)`[-.x]), na.rm = TRUE),
         `o_c_focus_Literacy (L)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Literacy (L)`[-.x]), na.rm = TRUE),
         `o_c_focus_Math (M)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Math (M)`[-.x]), na.rm = TRUE),
         `o_c_focus_Music and Movement (MM)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Music and Movement (MM)`[-.x]), na.rm = TRUE),
         `o_c_focus_None (N)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_None (N)`[-.x]), na.rm = TRUE),
         `o_c_focus_Other (O)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Other (O)`[-.x]), na.rm = TRUE),
         `o_c_focus_Other Language (OL)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Reading (R)`[-.x]), na.rm = TRUE),
         `o_c_focus_Science (SC)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Science (SC)`[-.x]), na.rm = TRUE),
         `o_c_focus_Social Studies (SS)_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_Social Studies (SS)`[-.x]), na.rm = TRUE),
         `o_c_focus_NA_classmean` = 
           map_dbl(row_number(), ~ mean(`o_c_focus_NA`[-.x]), na.rm = TRUE)) %>%
  ungroup 

#Re-format the teacher data so that it is one row per class
y2_teacher_obs <- y2_teacher_obs_raw %>% 
  dummy_cols(select_columns = c("o_t_verbal_o", "o_t_whom_o", "o_t_schedule_o", 
                                "o_t_task_o", "o_t_instruct", "o_t_focus_o", 
                                "o_t_tone_o", "o_t_attention_o", "o_t_es_o"),
             remove_selected_columns = TRUE) %>% 
    group_by(classid) %>%
  summarize(
    nsweeps = n(),
    nadult = length(unique(o_t_uniqueid)),
    across(starts_with(c("o_t_verbal_o", "o_t_whom_o", "o_t_schedule_o", 
                                "o_t_task_o", "o_t_instruct", "o_t_focus_o", 
                                "o_t_tone_o", "o_t_attention_o", "o_t_es_o")),
           mean)) %>%
  select(-ends_with("_")) %>% 
  ungroup

#Extract the caretype from the observation sheet
y2_coverpage_obs$

#Merge teacher and child observations 
y2_obs <- left_join(y2_child_obs, y2_teacher_obs, by = "classid")

```


Below we now input the child-level outcome data. We focus on the outcomes that Emily suggested, and extract the year 1 and year 2 values for each child and then merge to create a single dataset. 

```{r child.outcomes, warning=FALSE}
#Get Year 1 and Year 2 child data
y1_child_outcomes_raw <- read.dta13("y1c.dta")
y2_child_outcomes_raw <- read.dta13("y2c.dta")

#Rename all y1 variables and y2 variables so we don't lose them when merging
y1_child_outcomes <-  y1_child_outcomes_raw %>% 
  select(cid, c_mefs_str, c_pt_pcorrect, c_ltr_cogsoc_comp, c_ltr_emo_comp, 
         c_pra_total, c_pbsa_total, c_quils_total_raw, c_wjlw_str, c_wjap_str) %>% 
  rename_all( ~ paste0("y1_", .x)) %>% 
  mutate(cid = as.character(y1_cid))

y2_child_outcomes <-  y2_child_outcomes_raw %>% 
  select(cid, c_mefs_str, c_pt_pcorrect, c_ltr_cogsoc_comp, c_ltr_emo_comp, 
         c_pbsa_allgrades_total, c_pra_allgrades_total, c_quils_total_raw, 
         c_wjlw_str, c_wjap_str, c_age_cat_test, c_age_test) %>% 
  rename(c_pra_total = c_pra_allgrades_total,
         c_pbsa_total = c_pbsa_allgrades_total) %>% 
  rename_all( ~ paste0("y2_", .x)) %>% 
  mutate(cid = as.character(y2_cid))

#Merge Y2 and Y1 data together and calculate the gain score for each of the outcomes
child_outcomes <- merge(y1_child_outcomes, y2_child_outcomes, by = "cid") %>% 
  mutate(gain_c_mefs_str = y2_c_mefs_str - y1_c_mefs_str, 
         gain_c_pt_pcorrect = y2_c_pt_pcorrect - y1_c_pt_pcorrect,  
         gain_c_ltr_cogsoc_comp = y2_c_ltr_cogsoc_comp - y1_c_ltr_cogsoc_comp,  
         gain_c_ltr_emo_comp = y2_c_ltr_emo_comp - y1_c_ltr_emo_comp, 
         gain_c_pra_total = y2_c_pra_total - y1_c_pra_total, 
         gain_c_pbsa_total = y2_c_pbsa_total - y1_c_pbsa_total, 
         gain_c_quils_total_raw = y2_c_quils_total_raw - y1_c_quils_total_raw, 
         gain_c_wjlw_str = y2_c_wjlw_str - y1_c_wjlw_str, 
         gain_c_wjap_str = y2_c_wjap_str - y1_c_wjap_str,
         cid = as.numeric(cid))

```

Finally, we merge together the Year 1 observation data with the Year 1 & 2 child outcome data and add in care type. We omit observations that have no classroom observation resulting in a total analytic dataframe of 1169 observations of 64 variables. 

```{r merge.obs.outcomes}
#Merge in the outcomes data 
outcomes_and_obs <- left_join(child_outcomes, y2_obs, by = "cid") 

#Add in the care type
caretype <- read.dta13("y1caretype.dta") %>% 
  mutate(cid = as.character(cid))

#Remove observations that have no care type or no classroom observation
outcomes_and_obs <- left_join(outcomes_and_obs, caretype, by = "cid") %>% 
  mutate(hasobservation = is.na(classid)) %>% 
  filter(!is.na(caretype)) 

outcomes_and_obs_full <- outcomes_and_obs %>% 
  filter(!is.na(classid))

#Remove Y1 and Y2 data for cleanliness
outcomes_and_obs_full <- outcomes_and_obs_full %>% 
  select(-starts_with(c("y1", "y2")))

#Remove some irrelevant variables
outcomes_and_obs_full <- outcomes_and_obs_full %>% 
  select(-c(famid, dob:dob_uncertain, actual_fcc:hasobservation)) %>% 
  mutate(caretype = as.factor(caretype),
         actualtype = as.factor(actualtype))

#There are some issues with NA and NaN in the observation data that will mess up our analysis. We will replace these with zeros.
replace.na <- function(var){
  ifelse(is.na(var) | is.nan(var), mean(var), var)
}
names(outcomes_and_obs_full)
outcomes_and_obs_full <- outcomes_and_obs_full %>% 
  mutate_at(vars(o_c_verbal_talk:o_t_sched_t), replace.na)

dim(outcomes_and_obs_full)
```

## Analysis

We will first try a cross-validated LASSO, which will aggressively remove variables that do little to improve the predictive accuracy of the model.  

```{r mefs, echo=TRUE}
set.seed(4224)

#Subset to non-missing gain scores for the MEFS
gain_c_mefs_str_analysis <- outcomes_and_obs_full %>% 
  filter(!is.na(gain_c_mefs_str)) %>% 
  select(c(gain_c_mefs_str, o_c_verbal_talk:actualtype))

#Make the model matrix
x = model.matrix(gain_c_mefs_str ~ . , 
                 data=gain_c_mefs_str_analysis)
dim(x)
y = gain_c_mefs_str_analysis$gain_c_mefs_str
x = x[, -1]

# call cv.glmnet()
model_lasso <- cv.glmnet(x = x, y = y, alpha = 1)
plot( model_lasso)

model_lasso

# identify the best choice lambda (using the Minimum Lambda rule)
cc = coef( model_lasso, s = model_lasso$lambda.min)

# print out the model coefficients and store in a list.
cc = cc[cc[,1]!=0,1]
cc

```
This is not very satisfying! This is suggesting that *nothing* in our dataset is doing a particularly good job of predicting the gain in the MEFS score. Let's try with another variable.

```{r pt_pcorrect, echo=TRUE}
#Subset to non-missing gain scores for the MEFS
gain_c_pt_pcorrect_analysis <- outcomes_and_obs_full %>% 
  filter(!is.na(gain_c_pt_pcorrect)) %>% 
  select(c(gain_c_pt_pcorrect, o_c_verbal_talk:actualtype))

#Make the model matrix
x = model.matrix(gain_c_pt_pcorrect ~ . , 
                 data=gain_c_pt_pcorrect_analysis)
dim(x)
y = gain_c_pt_pcorrect_analysis$gain_c_pt_pcorrect
x = x[, -1]

# call cv.glmnet()
model_lasso <- cv.glmnet(x = x, y = y, alpha = 1)
plot( model_lasso)

# identify the best choice lambda (use the 1 SE rule)
cc = coef( model_lasso, s = model_lasso$lambda.min)

# print out the model coefficients and store in a list.
cc = cc[cc[,1]!=0,1]
sort(cc)

```

```{r rf_test, echo=TRUE, fig.height=8}

# Fit the random forest model
rf_fit <- train(gain_c_pt_pcorrect ~ ., #Use all variables in the prediction
                data = gain_c_pt_pcorrect_analysis, #Use the training data
                method = "ranger",
                importance = "permutation",
                num.trees = 500)

varImp(rf_fit) %>%
  pluck(1) %>%
  rownames_to_column("var") %>%
  ggplot(aes(x = reorder(var, Overall), y = Overall)) +
  geom_col(fill = "grey75") +
  coord_flip() +
  theme_minimal()

rf_fit <- train(gain_c_pt_pcorrect ~ ., #Use all variables in the prediction
rf_fit <- train(gain_c_pt_pcorrect ~ ., #Use all variables in the prediction
table(outcomes_and_obs_full$nclass)

#Next steps
#1. Try Y2 observation data instead of Y1, and compare results
#2. Try a model with a bunch of predictors from Y1 Y2
#3. Use the particular caretype from the  setting that is being observed in. (prov_type)
#Collapse CC-Community Based & CC- License Exempt
#4. Drop all variances and instead focus on means
#5. Incorporte child-level covariates 
#6. Age and elapsed time for assessment



```
